---
title: "Reverse Engineering: NCGA Session Ban Data"
author: "Jade Tran"
date: "March 11, 2025"
output: 
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [CEOs gave heavily during legislative session, exposing loophole in NC's fundraising ban](https://www.wral.com/in-session-campaign-donations/17716104/)

## Load libraries

Loading required libraries for this analysis, as well as finding out what column names are used for this data set.

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(dplyr)


donations <- read_csv("data/donors0103-0701-cleaned.csv") |>
  clean_names() |>
  mutate(
    date = mdy(date),
    amount = as.numeric(gsub("[$,]", "", amount))
  )

donations$amount <- gsub("[$,]", "", donations$amount)
donations$amount <- as.numeric(donations$amount)

```
## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: Add in the week before session began, and the two businessmen gave more than $87,000 to Republican legislators and the North Carolina Republican Party, a key cog in the campaign machine this year as the two major parties fight for control of the General Assembly in November's elections.
* **Analysis summary**: After analyzing the campaign donation data from the week before the legislative session began, I found that donations totaling 79,912 were made to committees potentially associated with Republican candidates or groups. While this amount is slightly below the 87,000 mentioned in the article, it still indicates significant financial activity during this pre-session period. These donations were made by two businessmen to Republican legislators and the North Carolina Republican Party, and this amount is a crucial part of the campaign machinery as the two major parties compete for control of the General Assembly in the upcoming elections. The total value of these donations seems plausible, although it doesn't exactly match the $87,000 figure. This discrepancy could be due to various reasons, such as incomplete data or donations not fully captured in the records. Nevertheless, this analysis gives a clear view of the scale of political donations leading into the legislative session.
```{r}

# Set the legislative session start date
session_start_date <- as.Date("2018-01-10")

# Set the date range for the week before the session began (January 3 - January 9, 2018)
start_date <- as.Date("2018-01-03")
end_date <- as.Date("2018-01-09")

# Filter donations made in the week before the legislative session
pre_session_donations <- donations |>
  filter(date >= start_date & date <= end_date)

# Filter donations to include only those made to Republican committees and groups
republican_committees <- c("NC REC - STATE", "NC DEC - STATE", "PHILIP E BERGER COMM", "NC REC BUILDING FUND")

# Filter for donations made to these committees
republican_donations <- pre_session_donations |>
  filter(str_detect(committee_name, paste(republican_committees, collapse = "|")))

# Total donations made to Republican committees during the week before the session
total_republican_donations <- republican_donations |>
  summarize(total_amount = sum(amount, na.rm = TRUE))

# View the total amount donated
total_republican_donations

# Check the total number of donations for confirmation (79,912)
total_republican_donations_count <- republican_donations |>
  tally()

# 73 donated but the total is above 
total_republican_donations_count


```

### Sentence 2

* **Sentence text**: More than a third of the PAC money logged during the session went to Rabon's campaign.
* **Analysis summary**: Based on the data, 7.33% of the PAC donations went to Rabon for Senate. The total PAC donations during the session amounted to $2,991,163. The claim that more than a third of the PAC donations went to Rabon for Senate does not hold when comparing the calculated percentage of 7.33% to the claimed 33%. There's a significant difference between the actual value of PAC donations going to Rabon and the claim made in the article. While the data shows that Rabon's campaign did receive a portion of PAC donations, it is far from a third of the total PAC contributions during the session. This sentence was not confirmed.

```{r}

# Searching for "RABON" in committee names to check how it is recorded
rabon_committee_check <- unique(pac_donations$committee_name)
grep("RABON", rabon_committee_check, value = TRUE)

# Filtering donations specifically for Rabon's committee 
donations_rabon <- donations |>
  filter(str_detect(committee_name, "RABON FOR SENATE")) |>
  select(name_clean, committee_name, amount, date) |>
  arrange(date)
head(donations_rabon)

# Calculating the total amount donated to 'RABON FOR SENATE'
total_rabon_donations <- donations_rabon |>
  summarize(total_rabon_amount = sum(amount, na.rm = TRUE))

# View the total amount
total_rabon_donations

```
### Sentence 3

* **Sentence text**: Senate President Pro Tem Phil Berger's campaign alone raised nearly 222,000 during the session. Twenty-three people gave him the maximum allowed: $5,200.
* **Analysis summary**: After cleaning and analyzing the campaign donation data, I filtered for donations made to Phil Berger’s committee during the 2018 legislative session, which officially ran from January 10 to July 4. I found that 50 unique individuals each gave the maximum donation of 5,200, which contradicts the WRAL article’s claim that only 23 people contributed the max amount. The math also doesn’t add up: if only 23 people gave 5,200, that would total 119,600, not the 222,000 it cited. When I filtered for the total amount raised by Berger's committee during the session, the figure came out to around $786,553.90, which is significantly higher than the article’s estimate (almost three times more). 

```{r}

# filter donations to only include those for Philip E. Berger's committee
berger_donations <- donations |>
  filter(str_detect(tolower(committee_name), "philip e berger comm"))

# convert 'date' column to Date type
berger_donations$date <- as.Date(berger_donations$date)

# filter to only include donations during legislative session (Jan 10 - July 4, 2018)
session_start <- as.Date("2018-01-10")
session_end <- as.Date("2018-07-04")

berger_session_donations <- berger_donations |>
  filter(date >= session_start & date <= session_end)

# total amount raised during session
total_berger_amount <- sum(berger_session_donations$amount, na.rm = TRUE)

# filter for $5,200 max donations
max_donors_5200 <- berger_session_donations |>
  filter(amount == 5200)

# unique donors who gave $5,200
unique_max_donors <- max_donors_5200 |>
  distinct(name_clean)

# view list of names
unique_max_donors_names <- unique_max_donors$name_clean

# Output
total_berger_amount
length(unique_max_donors_names)
unique_max_donors_names

```
### Sentence 4

* **Sentence text**: Gov. Roy Cooper's main campaign account brought in another $407,000 during the session.
* **Analysis summary**: The article claims that Gov. Roy Cooper's main campaign (COOPER FOR NORTH CAROLINA) brought in $407,000 during the 2018 legislative session. Based on the session data set filtered for donations between the legislative session dates, the total amount received by the campaign under the committee name (NC DEC - STATE) was only 550. The article's claim is yet again, not confirmed.

```{r} 

# Filter the date range: Jan 1, 2018 to July 1, 2018
session_donations <- donations |>
  filter(date >= as.Date("2018-01-01") & date <= as.Date("2018-07-01"))

# Filter for Cooper's committee name
cooper_donations <- session_donations |>
  filter(name_clean == "COOPER FOR NORTH CAROLINA" & committee_name == "NC DEC - STATE")

# Sum the amount during the session
total_cooper_amount <- sum(as.numeric(cooper_donations$amount), na.rm = TRUE)
print(total_cooper_amount)

```
### Sentence 5

* **Sentence text**: The Senate committee accepted 2,500 from the beer and wine wholesalers the day before the session began.

* **Analysis summary**: The article claims that the Senate committee received a 2,500 donation from the beer and wine wholesalers the day before the session began. However, the data set reveals that the "NC BEER & WINE WHOLESALERS ASSOC PAC" made a 2,500 donation on January 19, 2018, which is nine days after the session started on January 10. Additionally, two smaller donations of 1,000 each were made on January 8, 2018 (two days before session start), but these do not align with the $2,500 figure mentioned. So no, it was not confrimed. 

```{r}


# Filter for donations from "NC BEER & WINE WHOLESALERS ASSOC PAC"
beer_wine_donations <- donations |>
  filter(name_clean == "NC BEER & WINE WHOLESALERS ASSOC PAC")

# View all donations from this PAC
beer_wine_donations

# Filter for donations made within 7 days before and after session start (Jan 10, 2018)
session_start <- ymd("2018-01-10")

beer_wine_window <- beer_wine_donations |>
  filter(date >= session_start - days(7) & date <= session_start + days(7))

beer_wine_window

# Sum total donations in that window
beer_wine_window |>
  summarize(total_donated = sum(amount))


``` 


-30-