---
title: "lab_01"
author: "Derek Willis"
date: "2025-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your class reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the UMD courses data.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
umd_courses <- read_rds("lab_01.rds")
```

## Answer questions

**Q1** Which department has the highest mean (average) number of seats for its classes? You need to show the department with the highest average first. Which one is the most surprising/interesting to you? **A1** The Computer, Math and Natural Sciences department has rhe highest mean (average) number of seats for its classes. This isn't that surprising to me because that specific department is very high in demand, especially with CS becoming a big major at UMD. 

```{r}
umd_courses |>
  group_by(department) |>
  summarise(
    count_classes = n(),
    total_seats = sum(seats),
    mean_seats = mean(seats)
  ) |>
  arrange(desc(mean_seats))
```

**Q2** Find out the greatest number of times any individual class has been offered in our dataset. Your code must show the greatest number of classes first.
**A2** The greatest number of times any individual class has been offered in our dataset is a Doctoral Dissertation Research course at a whooping number of 1686 times.

```{r}
umd_courses |>
  group_by(title) |>
  summarise(times_offered = n()) |>
  arrange(desc(times_offered))

```

**Q3** Count the number of courses for each instructor and department. Which instructor has taught the most?
**A3** It may be my code, but the instructor is listed as NA. However, based on the data, I can see that it is an instructor of the Hearing and Speech Sciences.

```{r}
umd_courses |>
  group_by(instructors, department) |>
  summarise(courses_taught = n()) |>
  arrange(desc(courses_taught))
```

**Q4** Given the results for Q3, what would have made it easier to find the answer? You don't have to know *how* to accomplish this; I want you to describe in words what would make the answer easier to see. How confident are you that the answer you got is the right one, and why? **A4** It would have been easier to find the answer if the data set already came with entries within the data set aka instructor names rather than NA values. This could have been improved at the start of the data collection so then the answer wouldn't lend a wrong answer. I am not very confident that the answer I got is the wrong one because I didn't exactly get one but, if it is, it must be because of all the NA values under the instructor variable.

**Q5** Take one of the questions you listed in the final question of the pre_lab and try to answer it here in code. If you don't know how to fully answer it, write what code you can and describe in words what else you would need to do. If you got an answer, is it interesting? **A5** What department has the maximum seats? The department with the highest maximum seats is the University Courses department. I struggled with figuring out how to make the highest show first, so because it was showing lowest to highest first, I scrolled to the last page to see the highest which was 2876 maximum seats. I want to know how to combat this conflict with code. My code was probably wrong. This isn't that interesting though, instead it reads very vague. What department is University Courses? Aren't they all university courses?

```{r}
umd_courses |>
  group_by(department) |>
  summarise(
    count_classes = n(),
    total_seats = sum(seats),
    mean_seats = mean(seats),
    median_seats = median(seats),
    min_seats = min(seats),
    max_seats = max(seats)
  ) |>
  arrange(max_seats)
```
